Simulating Ultra High-Dimensional Multivariate Data Using the bigsimr R Package

### 2020-10-17 040-monte-carlo-evaluation.Rmd
cache computations and scale down until content finalized

```{r LoadLib040, echo = FALSE, eval = TRUE, message=FALSE, warning=FALSE, results = "hide"}
## devtools::install_github( "adknudson/bigsimr", ref = 'develop')
## knitr::opts_chunk$set(echo = FALSE, cache = TRUE)
## reticulate::use_condaenv("bigsimr")
library(bigsimr)
library(tidyverse)
## CORES <- as.integer( parallel::detectCores() - 1 )
set.seed(10162020)
```

@@

type <- c("pearson", "spearman", "kendall")
Cores <- as.integer( parallel::detectCores() - 1 )
n <- c(1e3, 1e4, 1e5)
adjustForDiscrete <- c(FALSE)

@@

```{r biNormPlot, echo = FALSE, eval = TRUE, message=FALSE, warning=FALSE, fig.height = 5, fig.width= 8, fig.align='center', fig.cap = "`bigsimr` recovers the Pearson specified correlations for MVN."}
dat <- readRDS("./results/norm_sims.rds")
dat %>%
    filter(cores == 1) %>%
    ggplot(aes(rho, rho_hat, color = type)) +
    ## ggplot(aes(rho, rho_hat)) +
    geom_point() +
    geom_abline(slope = 1) +
    ## facet_wrap(~ + N) + theme_bw()
    facet_wrap(~ type + N) + theme_bw()

>>>>

```{r LoadLib040, echo = FALSE, eval = TRUE, message=FALSE, warning=FALSE, results = "hide"}
## devtools::install_github( "adknudson/bigsimr", ref = 'develop')
knitr::opts_chunk$set(echo = FALSE, cache = TRUE)
## reticulate::use_condaenv("bigsimr")
library(bigsimr)
library(tidyverse)
cores <- as.integer( parallel::detectCores() - 1 )
set.seed(10162020)
```

@@

type <- c("pearson", "spearman", "kendall")
cores <- as.integer( parallel::detectCores() - 1 )
## n <- c(1e3, 1e4, 1e5)
n <- c(1e1, 1e2, 1e3)
adjustForDiscrete <- c(FALSE)

eps <- 1e-2
## grid_steps <- 100
grid_steps <- 10
sim_pars <- expand.grid(type = type, cores = cores, n = n,
                        stringsAsFactors = FALSE,
                        adjustForDiscrete = adjustForDiscrete)

@@

```{r biNormPlot, echo = FALSE, eval = TRUE, message=FALSE, warning=FALSE, fig.height = 5, fig.width= 8, fig.align='center', fig.cap = "`bigsimr` recovers the Pearson specified correlations for MVN."}
dat <- readRDS("./results/norm_sims.rds")
dat %>%
    filter(cores == cores) %>%
    ggplot(aes(rho, rho_hat, color = type)) +
    ## ggplot(aes(rho, rho_hat)) +
    geom_point() +
    geom_abline(slope = 1) +
    ## facet_wrap(~ + N) + theme_bw()
    facet_wrap(~ type + N) + theme_bw()

### 2020-10-17 050-motivating-application-and-examples.Rmd
cache computations and scale down until content finalized

```{r LoadLib050, echo = FALSE, eval = TRUE, message=FALSE, warning=FALSE, results = "hide"}
## devtools::install_github( "adknudson/bigsimr", ref = 'develop')
knitr::opts_chunk$set(echo = FALSE, cache = TRUE)
reticulate::use_condaenv("bigsimr")
library(bigsimr)
library(tidyverse)
CORES <- as.integer( parallel::detectCores() - 1 )
set.seed(10142020)
```

@@

```{r processBRCA, echo=FALSE, eval=TRUE, results='none'}
## full workflow simulating RNA-seq data
allDat <- readRDS( file = "~/Downloads/complete_processed_tcga2stat_RNASeq2_with_clinical.rds" )
lastClinical <- which( names(allDat) == 'tumorsize' )
brca <- allDat[ allDat$disease == "BRCA", (lastClinical+1):ncol(allDat) ]
dim(brca) ## num of genes 20501
## compute the median expression! avoid the 0s
brcaMedian <- apply(brca, 2, median)
## retain top (1-probs)*100% highest expressing genes for illustration
myProb <- 0.99
cutPoint <- quantile( x = brcaMedian, probs = myProb )
## genesToKeep <- names( brcaMedian ) [ which(brcaMedian >= cutPoint) ]
genesToKeep <- names( brcaMedian ) [ which(brcaMedian >= cutPoint) ]
brca <- brca[ , genesToKeep ]
## ncol(brca) / 20501
print(d <- length(genesToKeep))
## save data as rds
## saveRDS(brca, './data/brca99.rds')
```

```{r readBRCA, echo=FALSE, eval=FALSE}
## full workflow simulating RNA-seq data
brca <- readRDS(file = './data/brca99.rds') 
```

@@

```{r runBRCA, echo=TRUE, results=TRUE, cache = TRUE}
## Set the number of random vectors
n <- 10000
## construct margins
nb_margins <- make_nbinom_alist(sizes, probs)
## run sims
system.time( sim_nbinom <- rvec(n, nb_Rho, nb_margins, type = corType,
                                ensure_PSD = TRUE, cores = CORES) )
```

>>>>

```{r LoadLib050, echo = FALSE, eval = TRUE, message=FALSE, warning=FALSE, results = "hide"}
## devtools::install_github( "adknudson/bigsimr", ref = 'develop')
knitr::opts_chunk$set(echo = FALSE, cache = TRUE)
reticulate::use_condaenv("bigsimr")
library(bigsimr)
library(tidyverse)
cores <- as.integer( parallel::detectCores() - 1 )
set.seed(10142020)
```

@@

```{r processBRCA, echo=FALSE, eval=FALSE, results='none'}
## full workflow simulating RNA-seq data
allDat <- readRDS( file = "~/Downloads/complete_processed_tcga2stat_RNASeq2_with_clinical.rds" )
lastClinical <- which( names(allDat) == 'tumorsize' )
brca <- allDat[ allDat$disease == "BRCA", (lastClinical+1):ncol(allDat) ]
dim(brca) ## num of genes 20501
## compute the median expression! avoid the 0s
brcaMedian <- apply(brca, 2, median)
## retain top (1-probs)*100% highest expressing genes for illustration
myProb <- 0.99
cutPoint <- quantile( x = brcaMedian, probs = myProb )
## genesToKeep <- names( brcaMedian ) [ which(brcaMedian >= cutPoint) ]
genesToKeep <- names( brcaMedian ) [ which(brcaMedian >= cutPoint) ]
brca <- brca[ , genesToKeep ]
## ncol(brca) / 20501
print(d <- length(genesToKeep))
## save data as rds
saveRDS(brca, './data/brca99.rds')
```

```{r readBRCA, echo=FALSE, eval=TRUE}
## full workflow simulating RNA-seq data
brca <- readRDS(file = './data/brca99.rds') 
```

@@

```{r runBRCA, echo=TRUE, results=TRUE, cache = TRUE}
## Set the number of random vectors
n <- 10000
## construct margins
nb_margins <- make_nbinom_alist(sizes, probs)
## run sims
system.time( sim_nbinom <- rvec(n, nb_Rho, nb_margins, type = corType,
                                ensure_PSD = TRUE, cores = cores) )
```

### 2020-10-17 index.Rmd
cache computations and scale down until content finalized

```{r indexSetup, echo = FALSE, eval = TRUE, message=FALSE, warning=FALSE, results = "hide"}
library(tidyverse)
library(knitr)
library(rmarkdown)
library(bookdown)
## setwd('~/article_bigsimr')
## bookdown::render_book("index.Rmd", "bookdown::pdf_book")
## bookdown::serve_book()
set.seed(09212020)
```

@@

```{r serveHelper, include=FALSE, eval=FALSE}
## rmote::start_rmote()
## library(ggplot2); qplot(mpg, wt, data=mtcars, colour=cyl)
bookdown::serve_book()
```

>>>>

```{r indexSetup, echo = FALSE, eval = TRUE, message=FALSE, warning=FALSE, results = "hide"}
library(tidyverse)
library(knitr)
library(rmarkdown)
library(bookdown)
## setwd('~/article_bigsimr')
set.seed(09212020)
```

@@

```{r serveHelper, include=FALSE, eval=FALSE}
## rmote::start_rmote()
## library(ggplot2); qplot(mpg, wt, data=mtcars, colour=cyl)
## bookdown::render_book("index.Rmd", "bookdown::pdf_book")
bookdown::serve_book()
```
